Aquifer depletion exacerbates agricultural drought losses in the US High Plains

### 2023-11-03 Writing/nw_manuscript.Rnw
modify readme
minimum set of datasets added

All the raw data-sets used in this study are publicly accessible. The boundary shape file of the High Plains aquifer was obtained from the U.S. Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}. County and State boundaries were obtained using the R tigris package \citep{Rtigris}. These datasets were used for Figure \ref{fig:sat-map}. Corn and soybean yield data were obtained from the U.S. Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}. The R tidyUSDA package \citep{RtidyUSDA} was used to automatically download the relevant data. This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis. Weather data was obtained from the gridMET dataset \citep{Abatzoglou2013}, which is available at \url{https://www.climatologylab.org/gridmet.html}. Elevation data was obtained from the gridMET website at \url{https://www.climatologylab.org/gridmet.html}. These datasets were used to calculate water deficit, which was used for Figure \ref{fig:deficit-yield-hist} and the regression analysis. The aquifer thickness data were obtained from Hydroshare \citep{haacker2023} available at \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/}. This data is used for Figure \ref{fig:sat-map} and the regression analysis. Soil characteristics data was obtained from the Soil Survey Geographic Dataset, which was used for regression analysis. The R soilDB package \citep{Rsoildb} was used to download the data. Regression analyses are implemented using the R fixest package \citep{Rfixest}. Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations.

\subsection{Code Availability}
All data processing and analyses were carried out using R \citep{R}. The codes used for this study are available at the following GitHub repository: \url{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}.

@@

<<sat_ref, echo = FALSE>>=
sat_ref <-
  readRDS(here::here("Shared/Results/all_results.rds")) %>%
  .[1, sat_seq] %>%
  .[[1]] %>%
  .[length(.)]
@

>>>>

All the raw data-sets used in this study are publicly accessible. The boundary shape file of the High Plains aquifer was obtained from the U.S. Geological Survey at \url{https://water.usgs.gov/GIS/metadata/usgswrd/XML/ds543.xml#stdorder}. County and State boundaries were obtained using the R tigris package \citep{Rtigris}. These datasets were used for Figure \ref{fig:sat-map}. Corn and soybean yield data were obtained from the U.S. Department of Agriculture's National Agricultural Statistics Service (USDA-NASS), available at \url{https://www.nass.usda.gov/Quick_Stats/}. The R tidyUSDA package \citep{RtidyUSDA} was used to automatically download the relevant data. This data is used for Figure \ref{fig:deficit-yield-hist} and the regression analysis. Weather data was obtained from the gridMET dataset \citep{Abatzoglou2013}, which is available at \url{https://www.climatologylab.org/gridmet.html}. Elevation data was obtained from the gridMET website at \url{https://www.climatologylab.org/gridmet.html}. These datasets were used to calculate water deficit, which was used for Figure \ref{fig:deficit-yield-hist} and the regression analysis. The aquifer thickness data were obtained from Hydroshare \citep{haacker2023} available at \url{https://www.hydroshare.org/resource/7d925c7944244032af98c9ed20c22db6/}. This data is used for Figure \ref{fig:sat-map} and the regression analysis. Soil characteristics data was obtained from the Soil Survey Geographic Dataset, which was used for regression analysis. The R soilDB package \citep{Rsoildb} was used to download the data. Regression analyses are implemented using the R fixest package \citep{Rfixest}. Other key R packages include ggplot2 \citep{Rggplot2} for creating plots, data.table \citep{Rdatatable} for data wrangling, and sf \citep{Rsf} for spatial data operations. A minimum set of data files required to reproduce the results and figures presented in this article can be found on the Figshare website (DOI: 10.6084/m9.figshare.24492196). Due to file size constraints on Figshare, not every data file is hosted there. However, a complete set of data files can be accessed via the provided Dropbox link: \url{https://www.dropbox.com/scl/fo/bghhwlidmi7wx1ok0az5n/h?rlkey=tgbix1hp7g9np9etlo1z3biyr&dl=0}.

\subsection{Code Availability}
All data processing and analyses were carried out using R \citep{R}. The codes used for this study are available at the following GitHub repository: \url{https://github.com/tmieno2/Drought-Production-Risk-Aquifer}. Additionally, instructions for reproducing all the results and figures presented in this article can be found at the bottom of the repository page in the README section.

@@

<<sat_ref, echo = FALSE>>=
sat_ref <-
  readRDS(here::here("Results/all_results.rds")) %>%
  .[1, sat_seq] %>%
  .[[1]] %>%
  .[length(.)]
@
